Notes for ISIPTA poster

### 2016-06-13 asce-paper.tex
posternotes theory in

\author{
Gero Walter%
\thanks{
School of Industrial Engineering,
Eindhoven University of Technology, Eindhoven, The Netherlands.
E-mail: g.walter@tue.nl.},
\ Ph.D.
\\
and
Frank P.A. Coolen%
\thanks{
Department of Mathematical Sciences,
Durham University, Durham, United Kingdom.
E-mail: frank.coolen@durham.ac.uk.},
\ Ph.D.
}

@@

@@

\section{The Meat}

@@

@@

@@

>>>>

\author{
Gero Walter%
\thanks{
School of Industrial Engineering,
Eindhoven University of Technology, Eindhoven, The Netherlands.
E-mail: g.walter@tue.nl.},
\ Ph.D.
\\
and
Frank P.A. Coolen%
\thanks{
Department of Mathematical Sciences,
Durham University, Durham, United Kingdom.
E-mail: frank.coolen@durham.ac.uk.},
\ Ph.D.
\\
and ???
}

@@

Invited for special issue ``The treatment of uncertainty in risk and reliability modelling and decision making''
(special issue code SI022A), guest editors Luca Podolfillini, Bruno Sudret, Enrico Zio,
in ASCE-ASME Journal of Risk and Uncertainty in Engineering Systems: Part A (Civil Engineering)

@@

\section{Bayesian Analysis of Weibull Lifetimes}

@@

\section{Sets of Priors}

@@

***illustrate prior-data conflict similar to ESREL paper

\section{Robust RUL Estimation for Complex Systems via the Survival Signature}

\subsection{Right-censored Lifetimes}

@@

***change to general right-censoring? So $n_k-e_k \to c_k$, $e_k+c_k=n_k$,
introduce sets of observed and censored?

\subsection{Bayesian Estimation of Component Scale Parameter}

Posterior:
\begin{linenomath*}
\begin{align}
f_{\lambda_k\mid\ldots}(\lambda_k\mid\nkz,\ykz,\mbf{t}^k_{e_k;n_k})
 &\propto f_{\lambda_k}(\lambda_k)
          \big[ 1- F_k(\tnow\mid\lambda_k) \big]^{n_k-e_k}
          \prod_{i=1}^{e_k} f_k(t_i^k \mid \lambda_k) 
\end{align}
\end{linenomath*}
and so $\lambda_k\mid\nkz,\ykz,\mbf{t}^k_{e_k;n_k} \sim \ig(\nkn + 1, \nkn\ykn)$, where
\begin{linenomath*}
\begin{align}
\nkn + 1 &= \nkz + e_k + 1 \\
\nkn\ykn &= \nkz\ykz + (n_k-e_k) \tnow^\kappa + \sum_{i=1}^{e_k} (t_i^k)^\kappa
\end{align}
\end{linenomath*}
so the posterior expectation for $\lambda$ is
\begin{linenomath*}
\begin{align}
\E[\lambda_k\mid\nkn,\ykn] &= \ykn
                            = \frac{\nkz}{\nkz + e_k} \ykz +
                              \frac{e_k}{\nkz + e_k} \cdot \Big[\frac{n_k-e_k}{e_k} \tnow^\kappa + \frac{1}{e_k}\sum_{i=1}^{e_k} (t_i^k)^\kappa \Big]
\end{align}
\end{linenomath*}

\subsection{RUL using the Survival Signature}

System survival calculated by %(see Risk Analysis paper eq. (7))
\begin{linenomath*}
\begin{align}
P(\Tsys > t) &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \\
\intertext{ and assuming components of different types are independent,}
             &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K P(C^k_t = l_k)
\end{align}
\end{linenomath*}
where the survival signature $\Phi(l_1,\ldots,l_K)$ is non-decreasing in each $l_k$ for coherent systems,
and $P(C^k_t = l_k)$ is the (predictive) probability that exactly $l_k$ components of type $k$ function at time $t$.

A priori, we have, for $l_k = 0,1,\ldots, n_k$,
\begin{linenomath*}
\begin{align}
P(C^k_t = l_k\mid\nkz,\ykz)
 &= { n_k \choose l_k} \int [F_k(t \mid\lambda_k)]^{n_k-l_k}
                        [1 - F_k(t \mid\lambda_k)]^{l_k} f_{\lambda_k}(\lambda_k\mid\nkz,\ykz) \dd \lambda_k
\end{align}
\end{linenomath*}
under the assumption that components of the same type are independent given $\lambda_k$.

\subsection{Posterior Predictive Distribution}

For the situation with the system observed until $\tnow$, with data $\mbf{t}^k_{e_k;n_k}$,
we need the posterior predictive probability $P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$.

For $l_k > n_k - e_k$, we must have $P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k}) = 0$,
as at most $n_k - e_k$ components can function for $t > \tnow$.

%(Should one work with the subsystem consisting only of the non-failed $n_k - e_k$ components?)\\

A posteriori, we have, for $l_k = 0,1,\ldots, n_k - e_k$,
\begin{linenomath*}
\begin{align}
P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})
 &= { n_k - e_k \choose l_k} \int [P_k(T \leq t \mid T > \tnow, \lambda_k)]^{n_k - e_k - l_k} \times \\ & \hspace*{2ex}
                              [1 - P_k(T \leq t \mid T > \tnow, \lambda_k)]^{l_k}
    f_{\lambda_k\mid\ldots}(\lambda_k\mid\nkz,\ykz,\mbf{t}^k_{e_k;n_k}) \dd \lambda_k
\end{align}
\end{linenomath*}

Now,
\begin{linenomath*}
\begin{align}
P_k(T \leq t \mid T > \tnow, \lambda_k)
 &= \frac{P_k(\tnow < T \leq t \mid\lambda_k)}{P_k(T > \tnow \mid \lambda_k)}
  = \frac{F_k(t\mid\lambda_k) - F_k(\tnow\mid\lambda_k)}{1-F_k(\tnow\mid\lambda_k)} \\
 &= \frac{e^{-\frac{(\tnow)^\kappa}{\lambda_k}} - e^{-\frac{t^\kappa}{\lambda_k}}}{e^{-\frac{(\tnow)^\kappa}{\lambda_k}}}
  = 1 - e^{-\frac{t^\kappa - (\tnow)^\kappa}{\lambda_k}}
\end{align}
\end{linenomath*}

With this and the posterior substituted in, we get
\begin{linenomath*}
\begin{align}
\lefteqn{P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})} \\
 &= { n_k - e_k \choose l_k} \int \Big[1 - e^{-\frac{t^\kappa - (\tnow)^\kappa}{\lambda_k}}\Big]^{n_k - e_k - l_k}
                                  \Big[    e^{-\frac{t^\kappa - (\tnow)^\kappa}{\lambda_k}}\Big]^{l_k} \times \\ & \hspace*{28ex}
    \frac{[\nkn\ykn]^{\nkn + 1}}{\Gamma(\nkn+1)} \lambda_k^{-(\nkn + 1) - 1} e^{-\frac{\nkn\ykn}{\lambda_k}} \dd \lambda_k \\
 &= { n_k - e_k \choose l_k} \sum_{j=0}^{n_k-e_k-l_k} (-1)^j { n_k - e_k - l_k \choose j} \frac{[\nkn\ykn]^{\nkn + 1}}{\Gamma(\nkn+1)} 
    \times \\ & \hspace*{25ex}
    \int \lambda_k^{-(\nkn + 1) - 1} e^{-\frac{(l_k + j) (t^\kappa - (\tnow)^\kappa) + \nkn\ykn}{\lambda_k}} \dd \lambda_k
\end{align}
\end{linenomath*}

The terms remaining under the integral form the core of an $\ig(\nkn + 1, \nkn\ykn + (l_k + j) (t^\kappa - (\tnow)^\kappa))$,
allowing us to solve the integral using the corresponding normalization constant.
\begin{linenomath*}
\begin{align}
\lefteqn{P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})} \\
 &= { n_k - e_k \choose l_k} \sum_{j=0}^{n_k-e_k-l_k} (-1)^j { n_k - e_k - l_k \choose j}
    \left(\frac{\nkn\ykn}{\nkn\ykn + (l_k + j) (t^\kappa - (\tnow)^\kappa)}\right)^{\nkn + 1} \\
 &= \sum_{j=0}^{n_k-e_k-l_k} (-1)^j \frac{(n_k - e_k)!}{l_k! j! (n_k - e_k - l_k - j)!}   
    \left(\frac{\nkn\ykn}{\nkn\ykn + (l_k + j) (t^\kappa - (\tnow)^\kappa)}\right)^{\nkn + 1} \\
 &= \sum_{j=0}^{n_k-e_k-l_k} (-1)^j \frac{(n_k - e_k)!}{l_k! j! (n_k - e_k - l_k - j)!} \times \\ & \hspace*{13ex}  
    \left(\frac{\nkz\ykz + \sum_{i=1}^{e_k} (t_i^k)^\kappa + (n_k-e_k) (\tnow)^\kappa }%
               {\nkz\ykz + \sum_{i=1}^{e_k} (t_i^k)^\kappa + (n_k-e_k-l_k-j) (\tnow)^\kappa + (l_k + j) t^\kappa }\right)^{%
    \nkz + e_k + 1} 
\end{align}
\end{linenomath*}
for $l_k \in \{0,1,\ldots,n_k-e_k\}$.

We actually need to compute only those $P(C^k_t = l_k\mid \ldots)$ with $l_k$'s for which $\Phi(l_1,\ldots,l_K) > 0$.

Seen as function in $t$, we see that $t$ appears $n_k-e_k-l_k+1$ times,
once in each summand, in the denominator of the fraction powered to the $\nkn+1$,
and summands whith even $j$ are decreasing in $t$,
while summands whith odd $j$ are increasing in $t$.
Nevertheless, in total $P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$
must be decreasing in $t$, as the remaining $n_k-e_k$ components continue to age,
and so the probability that $l_k$ of them remain operational must decrease.

\subsection{Optimizing over Sets of Parameters}

%I would think that
%$\min / \max_{(\nkz,\ykz) \in \PkZ} P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$
%must be solved by numeric optimization for each $k, l_k, t$ 
%because of the alternating signs for the summands.
%(In absolute terms, each summand is increasing in $\ykz$, but not monotone in $\nkz$.)

%(Box-constraint optimization possible, e.g., with option \texttt{L-BFGS-B} of \texttt{optim} in \textbf{R}.)
Similar to the considerations about $t$, as higher values for $\ykz$ mean higher expected lifetimes for the components,
the larger $\ykz$, the higher the probability that many components survive
($P(C^k_t = l_k\mid\ldots)$ high for large $l_k$)
and, with that, the lower the probability that few or no components survive
($P(C^k_t = l_k\mid\ldots)$ low for small $l_k$).

In terms of the cumulative probability mass function (cmf) for $C^k_t$, which can be written as 
\begin{linenomath*}
\begin{align}
%F_{C^k_t}(l_k \mid \nkz,\ykz, \mbf{t}^k_{e_k;n_k}) = P(C^k_t \leq l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k}) 
% = \sum_{q=0}^{l_k} P(C^k_t = q\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})
F_{C^k_t}(l_k \mid \ldots) = P(C^k_t \leq l_k\mid\ldots) = \sum_{q=0}^{l_k} P(C^k_t = q\mid\ldots)\,,
\end{align}
\end{linenomath*}
we therefore conclude that
the lower cmf $\ul{F}(C^k_t = l_k\mid \ldots)$ must be obtained with the lower bounds for all $\ykz$, $k=1,\ldots,K$,
and the upper cmf $\ol{F}(C^k_t = l_k\mid \ldots)$ must be obtained with the respective upper bounds.

***proove rigorously (would need a kind of likelihood difference order)?
%Do I get your comments right that based on lower and upper bounds
%$\ul{P}(C^k_t = l_k\mid \ldots)$ and $\ol{P}(C^k_t = l_k\mid \ldots)$ for all $l_k = 0, 1, \ldots, n_k - e_k$,
%we can easily get lower and upper bounds for the cdf $F_{C^k_t}(l_k \mid \ldots)$?

Furthermore, higher expected lifetimes for the components must mean
a longer system survival, so in turn the $\ul{F}_{C^k_t}(l_k \mid \ldots)$'s
should give us the lower bound for $P(\Tsys > t\mid\ldots)$,
and the $\ol{F}_{C^k_t}(l_k \mid \ldots)$'s the upper bound.
Note that e.g.\ $\ul{F}_{C^k_t}(l_k \mid \ldots)$ may, however, not correspond to a single parameter pair $(\nkz, \ykz)$.
Also, in the system survival equation ($t > \tnow$)
\begin{linenomath*}
\begin{align} \label{eq:25}
\lefteqn{%
P(\Tsys > t\mid\{\nkz,\ykz, \mbf{t}^k_{e_k;n_k}\}_{k=1}^K)} \hspace*{10ex}\nonumber\\ 
 &= \sum_{l_1=0}^{n_1-e_1} \cdots \sum_{l_K=0}^{n_K-e_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K
    P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})
\end{align}
\end{linenomath*}
we cannot simply plug in all the $\ul{P}(C^k_t = l_k\mid \ldots)$ %and $\ol{P}(C^k_t = l_k\mid \ldots)$
to get $\ul{P}(\Tsys > t\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$, % and $\ol{P}(\Tsys > t\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$,
as the $\ul{P}(C^k_t = l_k\mid \ldots)$ could correspond to different $(\nkz,\ykz)$ for each $l_k$.
(It would give us a lower bound for $\ul{P}(\Tsys > t\mid\ldots)$, however, but it might be very coarse.)

Writing out Equation~\eqref{eq:25}, we get
\begin{linenomath*}
\begin{align}
\lefteqn{P(\Tsys > t\mid\{\nkz,\ykz, \mbf{t}^k_{e_k;n_k}\}_{k=1}^K)} \\
 &= \sum_{l_1=0}^{n_1-e_1} \cdots \sum_{l_K=0}^{n_K-e_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K
    P(C^k_t = l_k\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k}) \\
 &= \sum_{l_1=0}^{n_1-e_1} \cdots \sum_{l_K=0}^{n_K-e_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K
    \sum_{j=0}^{n_k-e_k-l_k} (-1)^j \frac{(n_k - e_k)!}{l_k! j! (n_k - e_k - l_k - j)!} \times \\ & \hspace*{45ex}
    \left(\frac{\nkn\ykn}{\nkn\ykn + (l_k + j) (t^\kappa - (\tnow)^\kappa)}\right)^{\nkn + 1} \\
 &= \sum_{l_1=0}^{n_1-e_1} \cdots \sum_{l_K=0}^{n_K-e_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K \label{eq:30}
    \sum_{j=0}^{n_k-e_k-l_k} (-1)^j \frac{(n_k - e_k)!}{l_k! j! (n_k - e_k - l_k - j)!} \times \\ & \hspace*{17ex}
    \left(\frac{\nkz\ykz + \sum_{i=1}^{e_k} (t_i^k)^\kappa + (n_k-e_k) (\tnow)^\kappa }%
               {\nkz\ykz + \sum_{i=1}^{e_k} (t_i^k)^\kappa + (n_k-e_k-l_k-j) (\tnow)^\kappa + (l_k + j) t^\kappa }\right)^{%
    \nkz + e_k + 1} 
\end{align}
\end{linenomath*}

While it is clear that the lower bounds for $\ykz$ will lead to the lower system survival function,
the role of $\nkz$ is not so clear.
We therefore resort to numerical optimization of \eqref{eq:25} over $\{\nkz, k=1,\ldots,K\}$
to obtain the lower bound for $P(\Tsys > t\mid\{\nkz,\ykz, \mbf{t}^k_{e_k;n_k}\}_{k=1}^K)$.
